---
title: "Final Project 131"
author: "Yuanning Li"
date: '2022-11-08'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 20)
```

## Introduction
 
The aim of the project is to analyze the population trend among the world from 1970s to 2020s, and research about which model can best fit on the population data and predict and population. I use data from $Kaggle$ and implement several techniques to answer the questions about population, growth rate, country areas, etc. I am also interested in fitting models on current data to predict the trend.

### Loading Packages and Setting Up The Environment

This project uses data from $Kaggle$ which records the population, land area, population density, etc. of 234 countries.

```{r, message=FALSE, warning=FALSE}
library(tidyverse)   # using tidyverse and tidymodels for this project mostly
library(tidymodels)
library(ggplot2)   # for most of our visualizations
library(rpart.plot)  # for visualizing trees
library(randomForest)   # for building our randomForest
```

```{r}
# import data
df <- read.csv('data/world_population.csv')
head(df)
```

### DATA DESCRIPTION

This database from $Kaggle$ contains 17 variables and 234 columns. While the codebook is provided in my text file, the variables listed here are useful for understanding this report.

`Rank`="Rank of Popluation"

`CCA3`="3 Digit Country/Territories Code"

`Country`="Name of the Country/ Territories"

`Capital`="Name of the Capital"

`Continent`="Name of the Continent"

`2022_Population`="Population of the Country/Territories in the year 2022"

`2020_Population`="Population of the Country/Territories in the year 2020"

`2015_Population`="Population of the Country/Territories in the year 2015"

`2010_Population`="Population of the Country/Territories in the year 2010"

`2000.Population`="Population of the Country/Territories in the year 2000"

`1990_Population`="Population of the Country/Territories in the year 1990"

`1980_Population`="Population of the Country/Territories in the year 1980"

`1970_Population`="Population of the Country/Territories in the year 1970"

`Area`="Area size of the Country/Territories in square kilometer"

`Density`="Population Density per square kilometer"

`Growth_Rate`="Population Growth Rate by Country/Territories"

`World_Population_Percentage`="The Population percentage by each Country/Territories"

Note: a full copy of the codebook is available in text files

### What is the current state of the world population? 
In June 2019, the world population estimate surveyed by the U.S. Census Bureau showed that the current global population is 757,713,040 people, which is much higher than the world population of 7.2 billion in 2015. It seems that the earth is very crowded and the population is still growing very fast.

### Why is analyzing and modeling the population so important?
For a growing population, since the resources on our planet are limited, the more the population, the greater the demand for various resources. When the world's population reaches a certain number, resources will be exhausted, and the loss of resources may lead to the destruction of the earth! So I hope these analysis and models can help us understand the current population situation and take corresponding measures to maintain the earth.

### Project pathway

Knowing the background and the importance of the topic, I'd like to discuss how to analyze and build models. First I do some initial data manipulation and cleaning on the original data. Then I explore the data and see if there's any interesting findings about population, density, area in time series and continent perspective. 

At the end, I use existing data from previous years to make the prediction. I'll split and resample the data, build the recipe and workflow, and train the model. Because this is a regression problem, I choose Ridge regression, Lasso regression, regression tree, and random forest these four models. I'd like to find the best parameters for our model using cross validation, and find which of the four models perform the best on our test dataset.

### Clean Data

Our original data is quite clean, so there's not much data cleaning to do. I simply renamed the feature names to make it more clean.

```{r}
# renamed all of the columns to make column names more clean and neat
colnames(df) <- c('Rank', 'CCA3', 'Country','Capital','Continent','Population_2022',
                  'Population_2020','Population_2015','Population_2010',
                  'Population_2000','Population_1990','Population_1980',
                  'Population_1970','Area', 'Density', 'Growth_Rate',
                  'World_Population_Percentage')
```

## Exploratory Data Analysis (EDA)

The entire exploratory data analysis will be based on the entire data set with 234 observations. Each observation represents the data of a country.

### Key Takeaways

##### -- What are the countries with largest and smallest population in the world? Is China going to remain the Top 1 population country in the world in the future?

##### -- Are the countries with small density / population has smaller growth rate than those countries with large density / population?

##### -- In the continent perspective, which continent has largest population and density? What is the trend?

##### -- What are the largest and smallest countries in size?

The EDA part will answer the above questions.

First extract the top five countries by population from 1970 to 2022. 
```{r}
# for EDA, no need for more dropping or cleaning, we will clean later for model
# add column for growth rate??
# EDA
# 1. Top 5 populous countries in 1970, 1980, 1990, 2000, 2010, 2015, 2020 and 2022.
df1 <- df
colnames(df1) <- c('Rank', 'CCA3', 'Country','Capital','Continent','2022',
                  '2020','2015','2010',
                  '2000','1990','1980',
                  '1970','Area', 'Density', 'Growth_Rate',
                  'World_Population_Percentage')

df1 <- df1 %>% select(c('Country','2022','2020',
                       '2015','2010',
                      '2000','1990',
                      '1980','1970')) %>% pivot_longer(cols=c('2022','2020',
                       '2015','2010',
                      '2000','1990',
                      '1980','1970'),
                      names_to = 'Year', values_to = 'Population') %>%
  arrange(desc(Population)) %>%
  group_by(Year) %>%
  slice(1:5)
head(df1, 10)
```

Then, use the above content to make a bar plot. it show the population of the top five populous countries from 1970 to 2022.

```{r}
ggplot(df1, aes(fill=Country, y=Population, x=Year)) + 
  geom_bar(position='dodge', stat='identity') +
  ggtitle ('Top 5 Populous Countries from 1970 to 2022')
```

From the figure, We can conclude that Russia's population growth is relatively small, so that it fell out of the top five after 1990. Brazil had a large population in 1990-2010, and it was not in the top five at other times. China has always maintained the first place, but India's population has grown relatively faster and has been catching up with China. In the very near future, India's population will soon catch up with China's.

Then, we extract the least five countries by population from 1970 to 2022.
```{r}
# 2. Top 5 populous countries in 1970, 1980, 1990, 2000, 2010, 2015, 2020 and 2022.
df2 <- df
colnames(df2) <- c('Rank', 'CCA3', 'Country','Capital','Continent','2022',
                  '2020','2015','2010',
                  '2000','1990','1980',
                  '1970','Area', 'Density', 'Growth_Rate',
                  'World_Population_Percentage')

df2 <- df2 %>% select(c('Country','2022','2020',
                       '2015','2010',
                      '2000','1990',
                      '1980','1970')) %>% pivot_longer(cols=c('2022','2020',
                       '2015','2010',
                      '2000','1990',
                      '1980','1970'),
                      names_to = 'Year', values_to = 'Population') %>%
  arrange(Population) %>%
  group_by(Year) %>%
  slice(1:5)
head(df2, 10)
```

The method I use to create a bar plot is similar to the previous question, it shows the population of the five least populous countries from 1970 to 2022.

```{r}
ggplot(df2, aes(fill=Country, y=Population, x=Year)) + 
  geom_bar(position='dodge', stat='identity') +
  ggtitle ('Top 5 Least Populous Countries from 1970 to 2022')
```

We can conclude that the population of Niue began to decrease from 1970 until 2010, and then increased a little. Montserrat is gaining momentum, starting in 2000 as the most populous of the five least populated countries, but Falkland Island has been growing as well.

Next, we look at the population growth rate for top 5 countries.

As described above, although China's population is currently the largest in the world, India is gaining momentum. Is it possible to surpass China and become the next country with the largest population in the world? 

```{r}
# 3. growth rate of the 5 most populous countries
df3 <- df1 %>% filter(Year == 2022)
df_tmp <- df %>% select(c('Country','Growth_Rate'))
df3 <- merge(df3, df_tmp, by='Country', all.x=TRUE) %>%
  select(c('Country','Population','Growth_Rate')) %>% arrange(desc(Population))
df3
```

From the data, the China has the least population growth rate and Pakistan has the highest population growth rate. Based on India's growth rate of 1.0068 is higher than China's growth rate of 1.0000, and the population until 2022, we expect India's population will have more than China's population.

Using the data `df3` obtained above, make a visualization to observe the population and growth trends of the top five countries in the world.

```{r}
ggplot(df3, aes(x=reorder(Country, +desc(Population))))  + 
  geom_bar(aes(y=Population*0.000000001, group=1),stat="identity", fill="cyan") +
  geom_line(aes(y=(Growth_Rate-1)*100, group=1),stat="identity",color="black",size=1)+
  scale_y_continuous(name = 'Population in Billion', sec.axis=sec_axis(trans = ~.*1, name="Increase Growth Rate in percent")) +
  labs(title= "Population vs. Growth Rate Top 5 Populous Countries", x = 'Country') +
  geom_point(aes(y=(Growth_Rate-1)*100, group=1), col='red') +
  geom_text(x=c(1,2,3,4,5), y = c(0.1, 0.8, 0.5, 0.8, 1.7), label=c(0, 0.68, 0.38, 0.64, 1.91))
```

To better visualize the growth rate, we changed population growth rate to increase of growth rate in percentage,for example, India's growth rate is 1.0068, so the increase of growth rate is 0.68%. 

From the plot, Pakistan has the highest increase of growth rate 1.91%. China has a increase of growth rate 0% in recent years, while India has a increase of growth rate 0.68%. If maintains, India's population will surpass that of China.

Then, we look at the comparison of population density. First look at the five countries with the lowest population density.
```{r}
# 4. The 5 countries with the least population density
df4 <- df %>% arrange(Density) %>% slice(1:5) %>%
  select(c("Country","Population_2022", "Density"))
df4
```

From the data, we know Greenland, Falkland Islands, Western Sahara, Mongolia and Namibia has the least population density.

Using the data `df4` obtained above, make a visualization to observe the population and density of the five countries with the least population in the world.
```{r}
ggplot(df4, aes(x=reorder(Country, +Population_2022)))  + 
  geom_bar(aes(y=Population_2022*0.000001, group=1),stat="identity", fill="blue") +
  geom_line(aes(y=Density, group=1),stat="identity",color="black",size=1)+
  scale_y_continuous(name = 'Population in Million', sec.axis=sec_axis(trans = ~.*1, name="Density (per km^2)")) +
  labs(title= "5 Least Density Countries", x= 'Country') +
  geom_point(aes(y=Density, group=1), col='red') +
  geom_text(x=c(1,2,3,4,5), y = c(0.51,0.23,2.36,3.31,2.37), label=c(0.31,0.03,2.16,3.11,2.17))
```

From the figure, the blue bars represents the population and the line represents the population density. Of the five countries, Namibia has the highest population density of 3.11$(people/km^2)$ and Greenland has the lowest of 0.03$(people/km^2)$.

Next, we look at the population growth rate for 5 countries with least density.

```{r}
# Growth rate of the 5 least densely populated countries
df5 <- df %>% arrange(Density) %>% slice(1:5) %>% 
  select(c("Country", "Density", "Growth_Rate"))
df5
```

From the data, the Greenland has the least population growth rate and Western Sahara has the highest population growth rate.

```{r}
ggplot(df5, aes(x=reorder(Country, +Density)))  + 
  geom_bar(aes(y=Density, group=1),stat="identity", fill="orange") +
  geom_line(aes(y=(Growth_Rate-1)*100, group=1),stat="identity",color="black",size=1)+
  scale_y_continuous(name = 'Density (per km^2)', sec.axis=sec_axis(trans = ~.*1, name="Increase Growth Rate in percent")) +
  labs(title= "5 Least Density Countries's Growth Rate", x = "Country") +
  geom_point(aes(y=(Growth_Rate-1)*100, group=1), col='red') +
  geom_text(x=c(1,2,3,4,5), y = c(0.6,0.6,2,1.7,1.7), label=c(0.4,0.43,1.81,1.51,1.46))
```

From the plot, the Western Sahara has the highest increase population growth rate of 1.81%, and the Greenland has the lowest increase population growth rate of 0.4%. comparing to the countries with large population and high density, the increase of growth rate of these low density countries are the similar. 


Then, Let's focus on the proportion of the population of different continents in the world from 1970 to 2022.

```{r}
# 6. Population distribution by continent
df6 <- df

df6 <- df6 %>% select(c('Continent','Population_2022','Population_2020',
                       'Population_2015','Population_2010',
                      'Population_2000','Population_1990',
                      'Population_1980','Population_1970')) %>% group_by(Continent) %>%
  summarise(sum_2022 = sum(Population_2022),
            sum_2020 = sum(Population_2020),
            sum_2015 = sum(Population_2015),
            sum_2010 = sum(Population_2010),
            sum_2000 = sum(Population_2000),
            sum_1990 = sum(Population_1990),
            sum_1980 = sum(Population_1980),
            sum_1970 = sum(Population_1970))

colnames(df6) <- c('Continent', '2022',
                  '2020','2015','2010',
                  '2000','1990','1980',
                  '1970')

df6 <- df6 %>% pivot_longer(cols=c('2022','2020',
                       '2015','2010',
                      '2000','1990',
                      '1980','1970'),
                      names_to = 'Year', values_to = 'Population') %>%
  arrange(Continent, Year)
df6[['Year']] <- as.numeric(df6[['Year']])
```

This time we use a line chart to do this visualization.

```{r}
ggplot(data = df6, aes(x=Year)) + 
  geom_line(aes(y=Population / 1000000000, colour=Continent)) + 
  geom_point(aes(y=Population / 1000000000, colour=Continent)) +
  labs(title= "Population distribution by Continent", y = "Population in Billion")
```

In the plot,Asia's population grows fast continuously form 1970s to 2022. And its population is much more than other continents.The other continent that has obvious population growth trend is Africa. Compare to Asia and Africa, other continents do not have a big growth throughout the years. The population of Oceania is the smallest of all continents, and the growth rate is also very small.

Then, Let's look at the population density of each of these continents.
```{r}
# 7. Population density of each continent

df7 <- df

df7 <- df7 %>% select(c('Continent','Population_2022','Population_2020',
                       'Population_2015','Population_2010',
                      'Population_2000','Population_1990',
                      'Population_1980','Population_1970', 'Area')) %>% group_by(Continent) %>%
  summarise(density_2022 = sum(Population_2022) / sum(Area),
            density_2020 = sum(Population_2020) / sum(Area),
            density_2015 = sum(Population_2015) / sum(Area),
            density_2010 = sum(Population_2010) / sum(Area),
            density_2000 = sum(Population_2000) / sum(Area),
            density_1990 = sum(Population_1990) / sum(Area),
            density_1980 = sum(Population_1980) / sum(Area),
            density_1970 = sum(Population_1970) / sum(Area))

colnames(df7) <- c('Continent', '2022',
                  '2020','2015','2010',
                  '2000','1990','1980',
                  '1970')

df7 <- df7 %>% pivot_longer(cols=c('2022','2020',
                       '2015','2010',
                      '2000','1990',
                      '1980','1970'),
                      names_to = 'Year', values_to = 'Density') %>%
  arrange(Continent, Year)
df7[['Year']] <- as.numeric(df7[['Year']])
```

Through the above calculation and classification, visualization can now be made.

```{r}
ggplot(data = df7, aes(x=Year)) + 
  geom_line(aes(y=Density, colour=Continent)) + 
  geom_point(aes(y=Density, colour=Continent)) +
  labs(title= "Population Density by Continent", y = "Density per km^2")
```

From the visualization, Asia is still far ahead of other continents. From a population density of about 65 people per square kilometer in 1970 to about 150 people per square kilometer in 2022. Oceania still has the lowest population density. In 1970, the population density was about 1 to 2 people per square kilometer. By 2022, about 3 to 4 people per square kilometer. 

By comparing population trend and population density trend from 1970 to 2020, we find the two trend shares the same pattern. 


Finally, we are interest in the land area of countries. The following distribution is the largest 5 countries and the smallest 5 countries.

```{r}
# Five largest countries
df8 <- df %>% arrange(desc(Area)) %>% slice(1:5) %>% select(c("Country", "Area"))
df8
```

The Largest country is Russia, the Area is 17098242 $km^2$.

```{r}
ggplot(df8, aes(x=reorder(Country, +desc(Area))))  + 
  geom_bar(aes(y=Area/1000000, group=1),stat="identity", fill="purple") +
  geom_text(x=c(1,2,3,4,5), y = c(16.098,10.985,10.707,10.373,9.516), label=c(17.098,9.985,9.707,9.373,8.516)) +
  labs(title= "Five Largest Countries", y = "Area in Million km^2", 
        x = 'Country')
```

In the visualization, Russia is shown to have the largest area, roughly twice the size of Brazil, which ranks fifth. The area gap between Canada, China and the United States, which ranks second to fourth, is relatively small.

```{r}
# Five smallest countries
df9 <- df %>% arrange(Area) %>% slice(1:5) %>% select(c("Country", "Area"))
df9
```

The smallest country is Vatican City, and it's area is 1$km^2$.

```{r}
ggplot(df9, aes(x=reorder(Country, +Area)))  + 
  geom_bar(aes(y=Area, group=1),stat="identity", fill="brown") +
  geom_text(x=c(1,2,3,4,5), y = c(2,3,7,13,20), label=c(1,2,6,12,21)) +
  labs(title= "Five Smallest Countries", y = "Area in Million km^2", 
        x = 'Country')
```

In the visualization, the Vatican is shown to have the smallest area at 1$km^2$, about the size of a square. The largest of them, Nauru, is only 21$km^2$.

## Data split & cross validation

For model training, we only keep the relevant features, including all the previous year's population, area, and continent. We delete features such as density and growth rate because they can be used to calculate population of 2022 directly (correlation is almost 1).  

The data was split in 70% training, 30% testing split. Stratified sampling was used as the `Continent`. 
```{r}
# keep only the relevant variables
data <- df %>% select(c(Continent, Area, Population_1970, 
                        Population_1980, Population_1990, Population_2000,
                        Population_2010, Population_2015, Population_2020,
                        Population_2022))
head(data) 
```

```{r}
# initial train test split
pop_split <- initial_split(data, strata = Continent, prop = 0.7)
pop_split
```

The training data set has 161 observations and the testing data set has 73 observations.

Then we use the cross validation resampling method to fold the training data into 10 folds with 5 repeats.
```{r}
pop_train <- training(pop_split)
pop_test <- testing(pop_split)
# 5 fold cross validation
pop_folds <- vfold_cv(pop_train, v = 5)
```

## Model Building 

Steps we use to build and analysis the model:

1. Build the recipe and workflow for each of the model.

2. Use cross validation to tune the model parameters, and find the best parameter for the model.

3. Use the best parameters found in step2, fit the model on test set and calculate model performance.

4. Compare the performance for each model. Find the best model in the four models.

### Building the Recipe

set up preprocess recipe for all models. 

```{r}
recipe <- 
  recipe(formula = Population_2022 ~ ., data = pop_train) %>% 
  step_novel(all_nominal_predictors()) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_zv(all_predictors()) %>% 
  step_normalize(all_predictors())
```

After many considerations, I decided to build the following four models.

1. Ridge Regression

2. Lasso Regression

3. Regression Tree

4. Random Forest

### Ridge Regression

Loaded the required object that I saved in my script, set `mode` to `"regression"`,tuned `penality`, and used the `glmnet` engine. I stored this model and `recipe` in workflow.

```{r}
set.seed(1234)
ridge_spec <- linear_reg(mixture = 0, penalty = tune()) %>%
  set_mode("regression") %>%
  set_engine("glmnet")

ridge_workflow <- workflow() %>% 
  add_recipe(recipe) %>% 
  add_model(ridge_spec)
```

Next, I set up the adjustment grid and updated the penalty parameters. I tune the penalty, in the range of $-10$ to $10$ with the level $50$. 

```{r}
ridge_penalty_grid <- grid_regular(penalty(range = c(-10, 10)), levels = 50)

ridge_tune_res <- tune_grid(
  ridge_workflow,
  resamples = pop_folds, 
  grid = ridge_penalty_grid)
```


### Lasso Regression

In a similar process, I set the model with tuning parameter penalty. Set the engine as glmnet and created a workflow.

The difference between Lasso Regression and Ridge Regression is Lasso regression 's mixture parameter equals 1 and Ridge regression's mixture parameter equals 0. This differs the two models from $L1$ to $L2$. 

```{r}
set.seed(1234)
lasso_spec <- 
  linear_reg(penalty = tune(), mixture = 1) %>% 
  set_mode("regression") %>% 
  set_engine("glmnet") 

lasso_workflow <- workflow() %>% 
  add_recipe(recipe) %>% 
  add_model(lasso_spec)
```

Next, set up the adjustment grid and updated the parameters. Tune the penalty, in the range of $-10$ to $10$ with the level $50$. 
```{r}
lasso_penalty_grid <- grid_regular(penalty(range = c(-10, 10)), levels = 50)

lasso_tune_res <- tune_grid(
  ridge_workflow,
  resamples = pop_folds, 
  grid = lasso_penalty_grid)
```


### Regression Tree

In this process, Set `mode` to `"regression"` and used the `rpart` engine. Stored this model and `recipe` in workflow.

```{r}
set.seed(1234)
reg_tree_spec <- decision_tree() %>%
  set_engine("rpart") %>%
  set_mode("regression")

reg_tree_wf <- workflow() %>%
  add_recipe(recipe) %>%
  add_model(reg_tree_spec %>% set_args(cost_complexity = tune()))
```

As above, Set up the adjustment grid and updated the parameters. Tune the cost complexity, in the range of -10 to -1 with the level 50. 

```{r}
reg_tree_param_grid <- grid_regular(cost_complexity(range = c(-10, -1)), levels = 50)

reg_tree_tune_res <- tune_grid(
  reg_tree_wf, 
  resamples = pop_folds, 
  grid = reg_tree_param_grid
)
```

### Random Forest

To prepare, load the required objects that I saved in my script, tuned `mtry`, where `mtry` is the number of levels of the trees. Set mode to `"regression"` , and used the random Forest engine. Stored this model and my recipe in a workflow.
```{r}
set.seed(1234)
rf_spec <- rand_forest(mtry = tune()) %>%
  set_engine("randomForest", importance = TRUE) %>%
  set_mode("regression")

rf_wf <- workflow() %>%
  add_recipe(recipe) %>%
  add_model(rf_spec)
```


Next, set up the tuning grid, and updated the parameters to tune the number of level of the trees. The minimum number of level is 1 and the maximum number of level is 9. 

```{r}
rf_grid <- grid_regular(parameters(rf_spec) %>%
 update(mtry = mtry(range= c(1, 9))), levels = 9)

rf_tune_res <- tune_grid(
  rf_wf, 
  resamples = pop_folds, 
  grid = rf_grid
)
```

### Repeated Cross Validation Parameter Tuning

### Ridge Regression

```{r}
autoplot(ridge_tune_res)
```

Use the `autoplot()`, we can observe the change of rmse and rsq when I tune the penalty of the ridge model. From the plot, we can see the rmse is small and stable, and rsq is large and stable at the beginning and has a sudden drop of rmse and sudden rise of rsq at the final. Therefore we pick the initial value as the best penalty value for the model.

```{r}
ridge_best <- select_best(ridge_tune_res, metric = "rsq")
ridge_best
```

Using the `select_best()` function, the best value of penalty is 1e-10.

### Lasso Regression

```{r}
autoplot(lasso_tune_res)
```

The trend of this graph is similar to the one above.

```{r}
lasso_best <- select_best(lasso_tune_res, metric = "rsq")
lasso_best
```

Also using the `select_best()` function, the best value of penalty is 1e-10.

### Regression Tree
```{r}
autoplot(reg_tree_tune_res)
```

For regression tree model, we tune the cost complexity parameter. From the plot, similar to ridge and lasso regression models, we can see the rmse is small and stable, and rsq is large and stable at the beginning and has a sudden drop of rmse and sudden rise of rsq at the final. Therefore we also pick the initial value as the best penalty value for the model.

```{r}
reg_tree_best <- select_best(reg_tree_tune_res, metric = "rmse")
reg_tree_best
```

As same before,  using the `select_best()` function, the best value of cost complexity is 1e-10.

### Random Forest

The autoplot shows when the number of levels of tree increase, the rmse decrease and the rsq increase. However, the best value of rmse appears at level 5 and best value of rsq appears at level 5.

```{r}
autoplot(rf_tune_res)
```

From the `autoplot()` we see the trend of rmse and rsq is getting better when we add layer to the trees. The best value of rmse and rsq appear when number of levels of trees is about 5. 

```{r}
rf_best <- select_best(rf_tune_res, metric = "rmse")
rf_best
```

Using the `show_best()` function, the best value of number of trees is config is 5 with $mtry = 5$.

## Model performance evaluation and select best model

Create a workflow with an adjusted name so I can identify it. Use the $fit()$ function for each model to run the models on the test set and find the best model by comparing r squared value.

### Final Ridge Regression Model

```{r}
ridge_final <- finalize_workflow(ridge_workflow, ridge_best)
ridge_final_fit <- fit(ridge_final, data = pop_train)
augment(ridge_final_fit, new_data = pop_test) %>%
  rsq(truth = Population_2022, estimate = .pred)
```

### Final Lasso Regression Model

```{r}
lasso_final <- finalize_workflow(lasso_workflow, lasso_best)
lasso_final_fit <- fit(lasso_final, data = pop_train)
augment(lasso_final_fit, new_data = pop_test) %>%
  rsq(truth = Population_2022, estimate = .pred)
```


### Final Regression Tree Model
```{r}
reg_tree_final <- finalize_workflow(reg_tree_wf, reg_tree_best)
reg_tree_final_fit <- fit(reg_tree_final, data = pop_train)
#reg_tree_final_fit %>%
  #extract_fit_engine() %>%
  #rpart.plot()
augment(reg_tree_final_fit, new_data = pop_test) %>%
  rsq(truth = Population_2022, estimate = .pred)
```


### Final Random Forest Model
```{r}
rf_final <- finalize_workflow(rf_wf, rf_best)
rf_final_fit <- fit(rf_final, data = pop_train)
augment(rf_final_fit, new_data = pop_test) %>%
  rsq(truth = Population_2022, estimate = .pred)
```

By comparison, Lasso Regression is the best model because it's r squared value is the largest. 

## Final model fitting and analysis

Compare the predictions of the four models across countries, actual values, and append them into a data frame for comparison.
```{r}
# append country, actual value, predict result from 4 models together in one data frame
data_pred_result <- df %>% select('Country', 'Population_2022')

data_pred_result['reg_tree_pred'] <- augment(reg_tree_final_fit, new_data = data)['.pred']

data_pred_result['rf_pred'] <- augment(rf_final_fit, new_data = data)['.pred']

data_pred_result['ridge_pred'] <- augment(ridge_final_fit, new_data = data)['.pred']

data_pred_result['lasso_pred'] <- augment(lasso_final_fit, new_data = data)['.pred']

data_pred_result
```

From the Table, In many countries, the model predictions are quite different from the actual values. For example, the actual value of American Samoa's population is 44273, but its Lasso Regression prediction result is 1507434; the actual population value of Anguilla is 15857, but its Lasso Regression prediction result is 1476729. These predictions are much larger than actual values, so the predictions are not accurate.

Advantage of lasso model: the value of its metric r squared value is the highest and closest to the actual value among the four models. In the other three models, there are negative numbers in the prediction results, but Lasso model's predict result stays positive in all its predicted values. 

## Conclusion

Through research, testing and analysis, Lasso Regression performed best in my models, Ridge Regression and Random forest are not bad, and Regression Tree did not perform well. Although the value of the r squared value of Lasso Regression, Random Forest and Ridge Regression is very close to 1, there is still a gap between the prediction results of many countries and their real population values. Because a lot of factors such as people’s life expectancy, fertility rate and death rate are all related to population growth, there is no way to predict the exact future world population.

In addition, the regression models has its limitation that it predicts a similar pattern for each continents or level of size of countries. However, from EDA at the beginning, even in the same continent or shares the similar value of areas, the country's growth rate differs. Therefore, for future research I propose to find more features to improve our model, such that gross national happiness index, and GDP, etc. We can also try more complicated models such as neuro network model. 

Overall, this World Population Modeling project provided me with a great opportunity to gain experience and improve my skills through data analysis and machine learning techniques.
